---
title: "Hierarchical Clustering - first"
author: "Chris coles"
date: "11 November 2015"
output: html_document
---

Prep the data frame for the clustering exercise. Straight forward plot show groups of points that should show up in the clustering.

The random variables are designed to produce obvious clusters. rnorm(12 - random numbers with a normal disrtibution and 12 off them. To cluster them, we set the standard deviation low at 0.2 and play with the mean value. So the first 4 values in y have a mean of 1 and and sd of 0.2 and we will produce 4 of them. The first 4 x values are also mean 1 and sd of 0.2  - these first 4 will be around 1,1 or in the lower left corner. The next set will be 2,2 and the last will be 1,3 - these arre the three groups visible in the plot. 

```{r}

set.seed(1234)
par(mar = c(0,0,0,0))
x <- rnorm(12, mean = rep(1:3, each = 4), sd = 0.2)
y <- rnorm(12, mean = rep(c(1,2,1), each = 4), sd = 0.2)
plot(x,y, col = "blue", pch = 19, cex = 2)
text(x + 0.05, y + 0.05, labels = as.character(1:12))
myFrame <- data.frame(x=x, y=y)
print(myFrame)
```

The next calculation is the distance between each point in the plot. This is carried out by using the `dist()` function. This will produce an object of class dist that has a table of distances between each point.

```{r}
distXY <- dist(myFrame)
print(distXY)

```

From this we can see that the distance between point 1 and point 2 is `r distXY[1]` and between point 7 and point 4 it is `r distXY[33]`. These distances are calculated using a the formula below with X and Y being the co ordinates of the poits. It is a imple equation that uses X and Y as two sides of a triangle and works out the direct line between the two points as the hyptonuse of the triangle - basic pythogourus theory. This is called the Euclidian distance. 

The formula for calculating the Euclidian distnce between two points is : 

$$\sqrt{\left(X_1 - X_2\right)^2 + \left(Y_1 - Y_2\right)^2} $$

The formula for calculating the Manhattan distance between two points is:

$$\sum_{i=1}^{n} |X_i - q_i| = |p_1 - q_1| + |p_2 - q_2| + |p_3 - q_3|...$$





